Finding the mass and occupation fraction of dark matter halos hosting Lyman-$\alpha$ emitters at $z\sim 3$: the importance of cosmic variance

### 2013-07-31 paper/ms.tex
conclusions

\title[Halo mass and occupation fraction for LAEs at
  $z=3.1$]{Mass and occupation
  fraction of dark matter halos hosting Lyman-$\alpha$ emitters
  at $z\sim 3$: limitations imposed by cosmic variance}    
\author[~J.~E. Forero-Romero and ~J.~E Mejia-Restreo]{
\parbox[t]{\textwidth}{\raggedright 
Jaime E. Forero-Romero$^{1}$ and
Julian E. Mej\'ia-Restrepo$^{2}$ 
}
\vspace*{6pt}\\
$^{1}$ Departamento de F\'{i}sica, Universidad de los Andes, Cra. 1
No. 18A-10, Edificio Ip, Bogot\'a, Colombia \\
$^{2}$ Departamento de Astronom\'{i}a, Universidad de Chile, Camino el
Observatorio 1515, Santiago, Chile}

@@

\begin{abstract}
%
We derive constraints on the mass and occupation fraction of dark
matter halos hosting \ly Emitting galaxies (LAEs) at a redshift of
$z=3.1$ by matching the number density and the angular
correlation function between mock and observed fields. We explicitly
take into account the cosmic variance on the typical observed field size by
constructing mock fields from a large cosmological N-body
simulation matching the observational geometries. To populate the
halos in the simulation we use a model where a dark matter halo with
mass in the range $M_{\rm   min}<M_{\rm h}<M_{\rm max}$ can only host one LAE with a
probability $f_{\rm occ}$. We find that the large scatter in the mocks, induced by
cosmic variance, makes the number density and spatial clustering
information insufficient to derive a unique and narrow set of values
for the parameters in the model. In particular, it is always possible
to find a model that is consistent with observations and has any
desired occupation fraction in the range $0.1\leq f_{\rm occ}\leq 1.0$
.  Nevertheless, in the models we can define three different families
of parameters based on their mass range $\Delta M \equiv
\log_{10}M_{\rm max} - \log_{10} M_{\rm min}$ and occupation fraction
$f_{\rm occ}$. The dominant family is composed by models with a narrow
mass range $\Delta M<1.0$ dex, a low occupation fraction $f_{\rm
  occ}\leq 0.3$ and a maximum mass $M_{\rm  max}< 10^{12}$\hMsun.  The
existence of this dominant family gives support to the idea that the
most massive dark matter halos at that epoch do not host the brightest
LAEs and that only a small fraction of star forming galaxies can be
actually detected as LAEs.

@@

One possible interpretation can be made in terms of a decreasing escape
fraction of \ly radiation in massive systems. There are detailed models for
radiative transfer that support the idea that massive galaxies with
higher metallicities have larger dust contents and less concentrated
ISM than lower mass systems which, due to the resonant nature of the
\ly line that increase the probability of absortion  of\ly photons, produce
high absorption of \ly photons but not of continuum or
other n-resonant lines \citep{ForeroRomero2011,Laursen2009}.

@@

In this \documentname we constrain the preferred mass
and occupation fraction of dark matter halos hosting Lyman Alpha Emitters at
redshift $z=3.1$ in a $\Lambda$CDM cosmology. We use a method that
matches the cosmic variance of the surface 
density number of LAEs between mock and real observations and is also
consistent with estimations of the angular
correlation function. The mock catalogs are constructed using a 
model with three basic parameters: the halo mass range where LAEs can
be found, $M_{\rm   min}<M_{\rm h}<M_{\rm   max}$, and the fraction of
the halos in this range that host a detectable LAE, $f_{\rm occ}$. After
a thorough exploration of the parameter space we find that a wide
variety of models are consistent with the observational constraints,
where only a minority has been considered so far in the
literature. Out of $9000$ initial combinations for the model
parameters we end up with $49$ successful arrangement of parameters.

The most important conclusion of this work is that available
observational information for the clustering properties of LAEs at
$z=3.1$ does not provide enough constraints to uniquely narrow the range of masses
and occupation fraction of dark matter halos hosting those
galaxies. In particular we are not able to put a stringent constraint
on the occupation fraction $f_{\rm occ}$.

In our results the models can be split into three families
depending the mass range $\Delta M=\log_{10}M_{\rm max} -
\log_{10}M_{\rm min}$ and the occupation fraction $f_{\rm occ}$. The first family is narrow both in
$\Delta M$ and $f_{\rm occ}$, a second family is only narrow in
$\Delta M$ and the third is narrow in $f_{\rm occ}$ and broad in
$\Delta M$.

All the halo mass ranges in the best models fall into the broad
range obtained in previous analysis \citep{Gawiser2007}. The
improvement of our method is being able to describe in high detail
all the possible models consistent with observations, including the
discovery of models previously overlooked like those with a high
occupation fraction $f_{\rm occ}>0.3$.

A central result in this exploration of parameter espace is the
existence of a large number of models with a very
narrow mass range $\Delta M< 0.1$ dex that are consistent in every
aspect with the spatial distribution of observed LAEs. In order to
have these characteristics it is required that halos around the mass
$M_{\rm max}$ for each model become inefficient in hosting detectable
galaxies. One way to achieve this is having LAEs with a decreasing \ly
escape fraction with increasing mass.

All the models also present a range of masses where the halos are
below the mass scale of $10^{11.4}$\hMsun \  inferred as a lower
threshold for the LBGs used in halo abundance matching
studies. Detailed study of high-z LAES will have a positive impact on
our understanding of the connection of low mass halos to star forming
galaxies.

@@

\section*{Acknowledgments} 
J.E.F-R thanks the hospitality of Changbom Park and the Korea
Institute for Advanced Study where the first full draft of this paper
was completed. J.E.F-R also thanks Peter Laursen for helpful comments
on the physical interpretation and the presentation of our results.

>>>>

\title[Halo mass and occupation fraction for LAEs at
  $z=3.1$]{Finding the mass and occupation
  fraction of dark matter halos hosting Lyman-$\alpha$ emitters
  at $z\sim 3$: the importance of cosmic variance}    
\author[~J.~E. Forero-Romero and ~J.~E Mejia-Restreo]{
\parbox[t]{\textwidth}{\raggedright 
Jaime E. Forero-Romero$^{1}$ and
Julian E. Mej\'ia-Restrepo$^{2}$ 
}
\vspace*{6pt}\\
$^{1}$ Departamento de F\'{i}sica, Universidad de los Andes, Cra. 1
No. 18A-10, Edificio Ip, Bogot\'a, Colombia \\
$^{2}$ Departamento de Astronom\'{i}a, Universidad de Chile, Camino el
Observatorio 1515, Santiago, Chile}

@@

\begin{abstract}
%
We derive constraints on the mass and occupation fraction of dark
matter halos hosting \ly Emitting galaxies (LAEs) at a redshift of
$z=3.1$ by matching the number density and the angular
correlation function between mock and observed fields. We explicitly
take into account the cosmic variance on the typical observed field size by
constructing mock fields from a large cosmological N-body
simulation matching the observational geometries. To populate the
halos in the simulation we use a model where a dark matter halo with
mass in the range $M_{\rm   min}<M_{\rm h}<M_{\rm max}$ can only host one LAE with a
probability $f_{\rm occ}$. We find that the large scatter in the mocks, induced by
cosmic variance, makes the number density and spatial clustering
information insufficient to derive a unique and narrow set of values
for the parameters in the model. In particular, it is always possible
to find a model that is consistent with observations and has any
desired occupation fraction in the range $0.1\leq f_{\rm occ}\leq 1.0$
.  Nevertheless, in the models we can define three different families
of parameters based on their mass range, $\Delta M \equiv
\log_{10}M_{\rm max} - \log_{10} M_{\rm min}$, and occupation fraction,
$f_{\rm occ}$. The dominant family is composed by models with a narrow
mass range $\Delta M<1.0$ dex, a low occupation fraction $f_{\rm
  occ}\leq 0.3$ and a maximum mass $M_{\rm  max}< 10^{12}$\hMsun.  The
existence of this dominant family gives support to the idea that the
most massive dark matter halos at that epoch do not host the brightest
LAEs and that only a small fraction of star forming galaxies can be
actually detected as LAEs.

@@

One possible interpretation can be made in terms of a decreasing escape
fraction of \ly radiation in massive systems. There are detailed models for
radiative transfer that support the idea that massive galaxies with
higher metallicities have larger dust contents and less concentrated
ISM than lower mass systems which, due to the resonant nature of the
\ly line that increase the probability of absortion  of\ly photons, produce
high absorption of \ly photons but not of continuum or
other n-resonant lines \citep{Laursen2009,ForeroRomero2011}.

@@

In this \documentname we look for constraints on the preferred mass
and occupation fraction of dark matter halos hosting Lyman Alpha Emitters at
redshift $z=3.1$ in a $\Lambda$CDM cosmology. We perform this study
paying attention to the impact of cosmic variance in these results. We
build a large number of mock catalogs matching observational
geometries. The mocks are constructed from a N-body simulation
following a simple recipe to assign a single LAE to each halo. Only
a fraction $f_{\rm occ}$ of halos with a mass range  $M_{\rm
  min}<M_{\rm h}<M_{\rm   max}$ can host a LAE. We proceed with a
thourough exploration of the space of free parameters $M_{\rm min},
M_{\rm max}$ and $f_{\rm occ}$ to find mocks that are consistent with
two observational constraints: the surface number density and the
angular correlation function. Out of the initial $9000$ initial
combinations of parameters in the model we find $49$ arrangements
consistent with observations.

The most relevant result from our study is the following. The 
successful parameter arrangements present a wide range of possibilities for
DM hosting LAE hosts at $z=3.1$. This means that the available LAEs' spatial
information is not sufficient to put a tight constraint on the mass
and occupation fraction of DM halos hosting LAEs. In particular it is
not possible to put a strong constraint on the occupation
fraction. Such a wide range in solutions is facilitated by the large
dispersion in the statistics derived from the mocks. All the halo mass
ranges and occupation fractions deduced in previous analysis in
previous analysis \citep[i.e.][]{Gawiser2007} are a subset of the
models we find in this \documentname.

Nevertheless, the sucessful models can be split into three families
depending the mass range $\Delta M=\log_{10}M_{\rm max} -
\log_{10}M_{\rm min}$ and the occupation fraction $f_{\rm occ}$. The
first family is narrow both in $\Delta M$ and $f_{\rm occ}$, a second
family is only narrow in $\Delta M$ and the third is narrow in $f_{\rm
  occ}$ and broad in $\Delta M$.

A second results from our investigation is the existence
of a large number of models with a very narrow mass range $\Delta M<
0.1$ dex that are consistent in every aspect with the spatial
distribution of observed LAEs. In order to have these characteristics
it is required that halos around the mass $M_{\rm   max}$ for each
model become inefficient in hosting detectable LAEs. Two possible
ways to achieve this is having LAEs with a decreasing \ly escape
fraction with increasing mass or increasing screening effects by
neutral Hydrogen around the most massive systems
\citep{Laursen2009,ForeroRomero2011}.

A third aspect that stands out is the existence of a dominant family
of models with narrow $\Delta M<0.1$, as described before, and also a
low occupation fraction $f_{\rm occ}<0.3$. This gives support to the
common wisdom stating that only a small fraction of star forming
galaxies can be detected as strong LAEs.

@@

\section*{Acknowledgments} 
J.E.F-R thanks the hospitality of Changbom Park and the Korea
Institute for Advanced Study where the first full draft of this paper
was completed. The authors also thank Peter Laursen, Paulina Lira and
Alvaro Orsi for helpful comments on the physical interpretation and
the presentation of our results.
